### 2021-04-04 Chapter-3/Chapter-3.tex
update

@@

  \begin{tabular}{llrrrrr}

@@

                       & \textbf{\# Instructions} & \textbf{Time}                                                               \\\hline

@@

    AES                & mbedTLS 2.5              & 68                & 0              & 68             & 39,855      & 512     \\
    AES                & mbedTLS 2.15             & 68                & 0              & 68             & 39,855      & 520     \\
    AES                & OpenSSL 0.9.7            & 75                & 0              & 75             & 1,704       & 231     \\
    AES                & OpenSSL 1.0.2f           & 88                & 0              & 88             & 1,350       & 36      \\
    AES                & OpenSSL 1.0.2k           & 88                & 0              & 88             & 1,350       & 35      \\
    AES                & OpenSSL 1.1.0f           & 88                & 0              & 88             & 1,420       & 36      \\
    AES                & OpenSSL 1.1.1            & 88                & 0              & 88             & 1,586       & 43      \\
    DES                & mbedTLS 2.5              & 15                & 0              & 15             & 4,596       & 58      \\
    DES                & mbedTLS 2.15             & 15                & 0              & 15             & 4,596       & 57      \\
    DES                & OpenSSL 0.9.7            & 6                 & 0              & 6              & 2,976       & 163     \\
    DES                & OpenSSL 1.0.2f           & 8                 & 0              & 8              & 2,593       & 166     \\
    DES                & OpenSSL 1.0.2k           & 8                 & 0              & 8              & 2,593       & 165     \\
    DES                & OpenSSL 1.1.0f           & 8                 & 0              & 8              & 4,260       & 182     \\
    DES                & OpenSSL 1.1.1            & 6                 & 0              & 6              & 8,272       & 229     \\

@@

    Chacha20           & Monocypher 3.0           & 0                 & 0              & 0              & 149,353     & 2       \\
    Poly1305           & Monocypher 3.0           & 0                 & 0              & 0              & 1,213,937   & 15      \\
    Argon2i            & Monocypher 3.0           & 0                 & 0              & 0              & 4,595,142   & 37      \\
    Ed25519            & Monocypher 3.0           & 0                 & 0              & 0              & 5,713,619   & 271     \\
    ECDSA              & mbedTLS 2.5              & 6                 & 6              & 0              & 4,214,946   & 48      \\
    ECDSA              & mbedTLS 2.15             & 4                 & 4              & 0              & 4,192,558   & 102     \\
    ECDSA              & OpenSSL 1.0.2f           & 5                 & 4              & 1              & 8,248,322   & 101     \\
    ECDSA              & OpenSSL 1.0.2k           & 5                 & 4              & 1              & 8,263,599   & 100     \\
    ECDSA              & OpenSSL 1.1.0f           & 5                 & 4              & 1              & 6,100,465   & 76      \\
    ECDSA              & OpenSSL 1.1.1            & 0                 & 0              & 0              & 10,244,076  & 121     \\
    ECDSA              & OpenSSL 1.1.1g           & 0                 & 0              & 0              & 9,266,191   & 102     \\

                       &                          &                   &                &                &             & minutes \\\cline{7-7}
    RSA                & mbedTLS 2.5              & 6                 & 6              & 0              & 22,109,246  & 39      \\
    RSA                & mbedTLS 2.15             & 12                & 12             & 0              & 24,484,441  & 44      \\
    RSA                & OpenSSL 0.9.7            & 107               & 105            & 2              & 17,002,523  & 23      \\
    RSA                & OpenSSL 1.0.2f           & 38                & 27             & 11             & 14,468,307  & 29      \\
    RSA                & OpenSSL 1.0.2k           & 36                & 27             & 9              & 15,285,210  & 40      \\
    RSA                & OpenSSL 1.1.0f           & 31                & 22             & 9              & 16,390,750  & 34      \\
    RSA                & OpenSSL 1.1.1            & 4                 & 4              & 0              & 18,207,016  & 8       \\
    RSA                & OpenSSL 1.1.1g           & 8                 & 8              & 0              & 18,536,796  & 5       \\
    RSA                & Libgcrypt 1.6.1          & 11                & 9              & 2              & 9,527,231   & 2       \\
    RSA                & Libgcrypt 1.7.3          & 14                & 14             & 0              & 10,513,606  & 14      \\
    RSA                & Libgcrypt 1.8.5          & 8                 & 8              & 0              & 27,407,986  & 113     \\

    Total              &                          & 904               & 241            & 663            & 167,141,947 & 341m    \\\hline
  \end{tabular}
\end{table*}

>>>>

We build the source program into 32-bit x86 Linux executables with GCC 7.5
running on Ubuntu 16.04.
We run our experiments on a 2.90GHz Intel Xeon(R) E5-2690 CPU with 128GB
RAM. The execution time is calculated on a single-core.

@@

  \begin{tabular}{llrrrrrr}

@@

                       & \textbf{\# Instructions} & \textbf{Time}     & \textbf{RAM (MB)}                                                  \\\hline

@@

    AES                & mbedTLS 2.5              & 68                & 0                 & 68             & 39,855      & 512     & 47    \\
    AES                & mbedTLS 2.15             & 68                & 0                 & 68             & 39,855      & 520     & 47    \\
    AES                & OpenSSL 0.9.7            & 75                & 0                 & 75             & 1,704       & 231     & 15    \\
    AES                & OpenSSL 1.0.2f           & 88                & 0                 & 88             & 1,350       & 36      & 15    \\
    AES                & OpenSSL 1.0.2k           & 88                & 0                 & 88             & 1,350       & 35      & 16    \\
    AES                & OpenSSL 1.1.0f           & 88                & 0                 & 88             & 1,420       & 36      & 16    \\
    AES                & OpenSSL 1.1.1            & 88                & 0                 & 88             & 1,586       & 43      & 16    \\
    DES                & mbedTLS 2.5              & 15                & 0                 & 15             & 4,596       & 58      & 8     \\
    DES                & mbedTLS 2.15             & 15                & 0                 & 15             & 4,596       & 57      & 8     \\
    DES                & OpenSSL 0.9.7            & 6                 & 0                 & 6              & 2,976       & 163     & 11    \\
    DES                & OpenSSL 1.0.2f           & 8                 & 0                 & 8              & 2,593       & 166     & 11    \\
    DES                & OpenSSL 1.0.2k           & 8                 & 0                 & 8              & 2,593       & 165     & 11    \\
    DES                & OpenSSL 1.1.0f           & 8                 & 0                 & 8              & 4,260       & 182     & 21    \\
    DES                & OpenSSL 1.1.1            & 6                 & 0                 & 6              & 8,272       & 229     & 43    \\

@@

    Chacha20           & Monocypher 3.0           & 0                 & 0                 & 0              & 149,353     & 2       & 15    \\
    Poly1305           & Monocypher 3.0           & 0                 & 0                 & 0              & 1,213,937   & 15      & 43    \\
    Argon2i            & Monocypher 3.0           & 0                 & 0                 & 0              & 4,595,142   & 37      & 102   \\
    Ed25519            & Monocypher 3.0           & 0                 & 0                 & 0              & 5,713,619   & 271     & 98    \\
    ECDSA              & mbedTLS 2.5              & 6                 & 6                 & 0              & 4,214,946   & 48      & 689   \\
    ECDSA              & mbedTLS 2.15             & 4                 & 4                 & 0              & 4,192,558   & 102     & 680   \\
    ECDSA              & OpenSSL 1.0.2f           & 5                 & 4                 & 1              & 8,248,322   & 101     & 980   \\
    ECDSA              & OpenSSL 1.0.2k           & 5                 & 4                 & 1              & 8,263,599   & 100     & 906   \\
    ECDSA              & OpenSSL 1.1.0f           & 5                 & 4                 & 1              & 6,100,465   & 76      & 705   \\
    ECDSA              & OpenSSL 1.1.1            & 0                 & 0                 & 0              & 10,244,076  & 121     & 1,048 \\
    ECDSA              & OpenSSL 1.1.1g           & 0                 & 0                 & 0              & 9,266,191   & 102     & 1,001 \\

                       &                          &                   &                   &                &             & minutes         \\\cline{7-7}
    RSA                & mbedTLS 2.5              & 6                 & 6                 & 0              & 22,109,246  & 39      & 3,708 \\
    RSA                & mbedTLS 2.15             & 12                & 12                & 0              & 24,484,441  & 44      & 4,012 \\
    RSA                & OpenSSL 0.9.7            & 107               & 105               & 2              & 17,002,523  & 23      & 3,601 \\
    RSA                & OpenSSL 1.0.2f           & 38                & 27                & 11             & 14,468,307  & 29      & 3,301 \\
    RSA                & OpenSSL 1.0.2k           & 36                & 27                & 9              & 15,285,210  & 40      & 3,402 \\
    RSA                & OpenSSL 1.1.0f           & 31                & 22                & 9              & 16,390,750  & 34      & 3,701 \\
    RSA                & OpenSSL 1.1.1            & 4                 & 4                 & 0              & 18,207,016  & 8       & 4,181 \\
    RSA                & OpenSSL 1.1.1g           & 8                 & 8                 & 0              & 18,536,796  & 5       & 3,901 \\
    RSA                & Libgcrypt 1.6.1          & 11                & 9                 & 2              & 9,527,231   & 2       & 2516  \\
    RSA                & Libgcrypt 1.7.3          & 14                & 14                & 0              & 10,513,606  & 14      & 2,701 \\
    RSA                & Libgcrypt 1.8.5          & 8                 & 8                 & 0              & 27,407,986  & 113     & 4,701 \\

    Total              &                          & 904               & 241               & 663            & 167,141,947 & 341m            \\\hline
  \end{tabular}
\end{table*}
Table~\ref{table:over_result} summarizes the results.
\tool{} finds 904 leaks in the crypto libraries.
Among these 904 leaks, 241 are due to secret-dependent
control-flow transfers and 663 are due to secret-dependent data accesses.

\tool{} also finds that most side-channel
vulnerabilities leak very little information in practice, which confirms our
initial assumption.
However, \tool{} finds some vulnerabilities with severe
leakages. Prior research has confirmed that some of these
vulnerabilities can be exploited in real attacks.
With our tool, developers can
distinguish non-critical ``vulnerabilities'' from severe ones.

Symmetric encryption implementations in OpenSSL and mbedTLS have significant leakage due to their lookup table implementations.
\tool{} confirms that those leakage comes from table lookups. The new implementation of OpenSSL has adopted several methods (e.g., one single S-box instead of four lookup tables, smaller lookup tables) to mitigate the problem. Those changes are rather easy but significantly decrease the total amount of leaked information as the quantification result indicates.

We also evaluate our tool on the RSA implementation. With the optimization
introduced in \S\ref{sec:scala}, we need not apply domain knowledge to
simplify the analysis. Our tool identifies all leakage sites
reported by CacheD~\cite{203878} and find new leaks in less time.
We also find newer
versions of RSA in OpenSSL have fewer leaks. We
will discuss the version changes and corresponding leakages in \S\ref{eval:asym}.

\subsection{Comparison with the Existing Tools}
\label{eval:scala}

In this section, we compare \tool{} with the
existing trace-based side-channel detection tools on vulnerability detections. For other tools, we use the data in the paper. As other tools do not quantify leakage sites,
we only include the time of detecting vulnerabilities to perform a fair comparison.

The comparison result with CacheD~\cite{203878} is shown in Table~\ref{eval:cacheD}.
Note that one statement in the source code can be compiled into several machine instructions. So it is possible that one statement can have multiple leakage points. Under the circumstance, we think it is only a leakage.
We have confirmed that \tool{} can identify all the secret-dependent data access vulnerabilities reported by CacheD. In addition, \tool{} finds many new ones.
CacheD fails to detect some vulnerabilities for two
reasons. First, CacheD can only detect secret-dependent data access
vulnerabilities. \tool{} can detect secret-dependent control-flows as well.
Second, according to the CacheD paper, CacheD times out after 20 hours to process
asymmetric ciphers. CacheD applies some domain knowledge to simplify and speed up
the analysis.
While those optimizations do not introduce false positives, they may miss some
vulnerabilities.
We notice that the number of instructions in those traces are different due to the different analysis starting
functions and building options during the evaluation.
Table~\ref{eval:cacheD} shows that
\tool{} is faster than CacheD. \tool{} is much faster than CacheD when analyzing the same
number of instructions. For example, when we test~\tool{} on AES from OpenSSL
0.9.7, \tool{} is over 100x faster than CacheD.
\begin{table}[]
  \caption{Comparison with CacheD: Time,
    Secret-dependent Control-flows\,(CF), Secret-dependent Data-flows\,(DF), The Number of Instructions\,(\# Instructions).}
  \label{eval:cacheD}
  \begin{threeparttable}
    \begin{tabular}{@{}c|r@{~~}r@{~~}r|r@{~~}r@{~}r@{}}
      \multicolumn{1}{l|}{} & \multicolumn{3}{c|}{CacheD} & \multicolumn{3}{c}{Abacus}                                             \\
      Name                  & Time (s)                    & \# Instructions            & DF & Time (s) & \# Instructions & CF+DF   \\ \hline
      AES\tnote{1}          & 43.4                        & 791                        & 48 & 0.23     & 1,704           & 0+75    \\
      AES\tnote{2}          & 48.5                        & 2,410                      & 32 & 0.04     & 1,350           & 0+88    \\
      RSA\tnote{1}          & 199.3                       & 674,797                    & 1  & 1,351.6  & 17,002,523      & 105+2   \\
      RSA\tnote{2}          & 165.6                       & 473,392                    & 1  & 1,753.3  & 14,468,307      & 27+11   \\
      RSA\tnote{3}          & 11542.3                     & 26,848,103                 & 2  & 128.1    & 9,527,231       & 9+2     \\
      RSA\tnote{4}          & 10788.9                     & 27,775,053                 & 0  & 891.7    & 10,513,606      & 14+0    \\ \hline
      Total                 & 22,788.0                    & 55,738,546                 & 84 & 4,125.0  & 51,514,721      & 155+178 \\ \hline
      \multicolumn{7}{l}{\# of Instructions per second \qquad  CacheD: 2,445 \qquad \tool: 12,489}                                 \\ \hline
    \end{tabular}
  \end{threeparttable}
  \begin{tablenotes}
    \scriptsize
    \item[1] OpenSSL\,0.9.7  \item[2] OpenSSL\,1.0.2f   \item[3] Libgcrypt\,1.6.1 \item[4] Libgcrypt\,1.7.3 \\
  \end{tablenotes}
  \vspace*{-20pt}
\end{table}
Since DATA~\cite{217537} needs to compare several execution
traces to identify side-channel leakages, \tool{} also outperforms
DATA in terms of performance. For
example, it takes 234 minutes for DATA to analyze the RSA of
implementation in OpenSSL 1.1.0f. \tool{} only spends 34 minutes
according to Table~\ref{table:over_result}.  Also, DATA reports
report 278 control-flow and 460 memory-access leaks. Among those
leakages, they find one new vulnerability in RSA after some manual
analysis.  \tool{} finds the vulnerability and reports the
vulnerability is severe (\textsf{int\_bn\_mod\_inverse} leaks more
than 14.9 bits and \textsf{BN\_div} leaks more than 17.2 bits),
which eases the pain to identify real sensitive leaks.}{DATA~\cite{217537} identifies
side-channel leakages by finding differences in execution traces of
the test program under \emph{various secret inputs}. According to
the original DATA paper, it uses 443 different traces to analyze the
side-channel vulnerabilities in symmetric cyphers and 450 different
traces to analyze the side-channel vulnerabilities in asymmetric
cyphers. On the other hand, \tool{} detects side-channel
vulnerabilities from one execution trace. \tool{} uses symbolic
analysis to extract formulas that model each side-channel
leakage. After that, we sample the formula with \emph{various
  secret inputs} to detect and quantify each leakage site. In
theory, DATA might have better code coverage than \tool{} because it
uses more execution traces,  but \tool{} has the following
advantages. a) \tool{} is faster than DATA. For example, it takes
116 minutes for DATA to detect vulnerabilities in the RSA implementation in OpenSSL 1.1.0f\@. \tool{} only spends 34 minutes, as shown in Table~\ref{table:over_result}. It takes 13 minutes and 20 minutes
for DATA to analyze the side-channel leakages in AES and DES,
respectively. On the other hand, \tool{} finishes its analysis in
less than ten seconds while finding all the leakages reported by
DATA. b) Because \tool{} does not run the test program again when we
have a new \emph{secret input}, \tool{} can test more input secrets
on those formulas within the same time to achieve better precision.
c) DATA tries to use leakage models (domain knowledge) to classify each leakage.
The strength of Abacus is that it does not need such domain knowledge.
DATA reports 278 control-flow and 460 memory-access leaks for the RSA implementation in OpenSSL 1.1.0f. Among those leakages, they find one new vulnerability in RSA after some manual analysis.
\tool{} finds the vulnerability and reports the vulnerability is
severe (\textsf{int\_bn\_mod\_inverse} leaks more than 14.9 bits and
\textsf{BN\_div} leaks more than 17.2 bits), which helps identify
real sensitive leaks.  For each leakage site, \tool{} can provide
concrete examples to trigger the issue and give an estimation to
assess the severity level of the vulnerability.
